Placing the spotted T Tauri star LkCa 4 on an HR diagram

### 2016-08-16 document/ms.tex
This is version 0.6 of the paper, the one we will send to co-authors!

\begin{figure*}
 \centering
 \includegraphics[trim=1.6cm 12.6cm 2.2cm 2.4cm, clip=true, width=0.60\textwidth]{figures/vband_spot_2750k}
\caption{The $V$-band magnitude in 2014--2015, converted into fill factor for the cool component.  The optical brightness depends mostly on the hot component.  If we fix a 75\% filling factor, as measured in the IGRINS spectrum, to $V=12.83$ at the time of the observation, then the $V$-band amplitude  corresponds to filling factors of 67--83\%.  The factor of $\sim 2$ change in visible surface area of the hot component, from 33\% to 17\%, is required to produce the $\Delta V=0.6$. \todo{Update with all available photometry or the derived multi-term fit} }
\label{fig:vband_spot}
\end{figure*}

@@

\todo{Some overview?}

@@

An accurate extinction estimate from the full SED is challenging because of a strong dependence on the areal distribution of temperatures in the photosphere, which in our work is simplified to a two temperature fit.  However, isolating specific colors constrains the possible range of extinctions.  
The $B-V$ color is stable and not affected by cooler components, and may therefore be compared with the expected colors for the hotter (4100 K) temperature.  
The average $B-V=1.41$ mag color from \citet{grankin08} has a color excess $E(B-V)=0.09$ mag relative to the empirical color expected for a main sequence star and $E(B-V)=0.20$ mag for colors of pre-main sequence stars, based on the temperature-color tables of \citep{pecaut13}.  The main sequence colors are probably most 
appropriate in this analysis because they are more free of spots.  
The $B-V$ color 
excess leads to an extinction $A_V=0.28\pm0.15$ mag, or $0.62\pm0.15$ mag relative to 
pre-main sequence colors.   The listed uncertainties are calculated from extinction estimates relative to colors of 4000 and 4200 K photospheres.
The $V-R_J$ color excess leads to $A_V=0.48$ mag, which should be less affected by gravity (based on $V-I_C$ colors in Pecaut et al.) but is likely to be overestimated because the cool component contributes to the $R_J$-band emission, as seen in Fig.~\ref{fig:colors}.
In the near-IR, the estimated $V=12.61$ mag at the time of the 2MASS observations leads to a template (main sequence) color 
$J-K_s=0.87$ mag, with a range of $\sim 0.05$ mag for acceptable solutions with other 
temperatures and covering fractions. The observed $J-K=0.93$ mag leads to $A_V=0.35$ mag, 
with a range of $0.29$ mag to account for the range of other acceptable photospheric solutions.  
The template near-IR color is more robust than optical colors to differences 
in the main sequence/pre-main sequence temperature scales because the pre-main 
sequence $J-K$ color is redder at 4100 K and bluer at 2750 K {\bf GJH: (just redo for Pec13 young colors)}.  Based on this 
analysis, an $A_V=0.3$ mag is adopted throughout this paper.

The near-IR extinctions of LkCa 4, and perhaps other TTSs, will be overestimated if emission is generated from a multi-temperature spectrum but template colors are based on only the hotter component.  This mismatch likely explains why near-IR extinctions are often higher than optical extinctions for young stars \citep[see analysis in][]{herczeg14}.  Our results tentatively support the lower extinctions obtained at optical wavelengths, although the template colors used to calculate extinction are uncertain and will need revision.  Differences between main sequence and pre-main sequence temperature scales and colors in the \citet{pecaut13} tables are introduced by a combination of spots, in addition to any color differences caused by gravity differences.
Uncertainty in colors may be solved with self-consistent comparisons to young stars with similar spectral types, if spot coverage is also similar.

@@

{\bf For the conclusions}

Discrepancies in colors, spectral type/effective temperatures, and extinctions for LkCa 4, and perhaps more broadly for TTSs, are dominated by ignoring spots and are not caused by fundamental problems in either photospheric models or in spectral type-effective temperature scales.

{\bf GJH TO DO:  revise plots;  TiO Table; more citations for colors}

0 New

Reply

@@

One conceivable interpretation of the 67\% to 83\% coverage fraction of cool photosphere is a stellar axis inclined towards our observing direction, with a large circum-polar spot covering much of the observable hemisphere, as described by \citet{donati14}.  The ambient hot photosphere would then be close to the equator, which would be limb-darkened due to projection effects.  Such limb darkening would cause a non-linear transformation between our observed fill factor $f_{\Omega}$, and the surface areal coverage fraction of spots.  So it's conceivable, and probably likely that the reported $f_{\Omega}$ is and overestimate of the surface area actually covered by spots.  In this circum-polar spot scenario, local fingers of cool spots could reach down to near the equator, entering into and out-of the field of view and giving rise to the large amplitude photometric variation seen on LkCa 4.  \todo{Donati-- How does this compare to the reconstructed maps from ZDI in Donati et al. 2014?}.

@@

@@

\todo{Verify that the final term possesses a $q$ and not a $q^2$ term.}

>>>>

\begin{figure*}
 \centering
 \includegraphics[trim=1.6cm 12.6cm 2.2cm 2.4cm, clip=true, width=0.60\textwidth]{figures/vband_spot_2750k}
\caption{The $V$-band magnitude in 2014--2015, converted into fill factor for the cool component.  The optical brightness depends mostly on the hot component.  If we fix a 75\% filling factor, as measured in the IGRINS spectrum, to $V=12.83$ at the time of the observation, then the $V$-band amplitude  corresponds to filling factors of 67--83\%.  The factor of $\sim 2$ change in visible surface area of the hot component, from 33\% to 17\%, is required to produce the $\Delta V=0.6$. }
\label{fig:vband_spot}
\end{figure*}
%\todo{Update with all available photometry or the derived multi-term fit}

@@

%\todo{Some overview?}

@@

An accurate extinction estimate from the full SED is challenging because of a strong dependence on the areal distribution of temperatures in the photosphere, which in our work is simplified to a two temperature fit.  However, isolating specific colors constrains the possible range of extinctions.  
The $B-V$ color is stable and not affected by cooler components, and may therefore be compared with the expected colors for the hotter (4100 K) temperature.  
The average $B-V=1.41$ mag color from \citet{grankin08} has a color excess $E(B-V)=0.09$ mag relative to the empirical color expected for a main sequence star and $E(B-V)=0.20$ mag for colors of pre-main sequence stars, based on the temperature-color tables of \citep{pecaut13}.  The main sequence colors are probably most 
appropriate in this analysis because they are more free of spots.  
The $B-V$ color 
excess leads to an extinction $A_V=0.28\pm0.15$ mag, or $0.62\pm0.15$ mag relative to 
pre-main sequence colors.   The listed uncertainties are calculated from extinction estimates relative to colors of 4000 and 4200 K photospheres.
The $V-R_J$ color excess leads to $A_V=0.48$ mag, which should be less affected by gravity (based on $V-I_C$ colors in Pecaut et al.) but is likely to be overestimated because the cool component contributes to the $R_J$-band emission, as seen in Fig.~\ref{fig:colors}.
In the near-IR, the estimated $V=12.61$ mag at the time of the 2MASS observations leads to a template (main sequence) color 
$J-K_s=0.87$ mag, with a range of $\sim 0.05$ mag for acceptable solutions with other 
temperatures and covering fractions. The observed $J-K=0.93$ mag leads to $A_V=0.35$ mag, 
with a range of $0.29$ mag to account for the range of other acceptable photospheric solutions.  
The template near-IR color is more robust than optical colors to differences 
in the main sequence/pre-main sequence temperature scales because the pre-main 
sequence $J-K$ color is redder at 4100 K and bluer at 2750 K.  Based on this 
analysis, an $A_V=0.3$ mag is adopted throughout this paper.

The near-IR extinctions of LkCa 4, and perhaps other TTSs, will be overestimated if emission is generated from a multi-temperature spectrum but template colors are based on only the hotter component.  This mismatch likely explains why near-IR extinctions are often higher than optical extinctions for young stars \citep[see analysis in][]{herczeg14}.  Our results tentatively support the lower extinctions obtained at optical wavelengths, although the template colors used to calculate extinction are uncertain and will need revision.  Differences between main sequence and pre-main sequence temperature scales and colors in the \citet{pecaut13} tables are introduced by a combination of spots, in addition to any color differences caused by gravity differences.
Uncertainty in colors may be solved with self-consistent comparisons to young stars with similar spectral types, if spot coverage is also similar.  Discrepancies in colors, spectral type/effective temperatures, and extinctions for LkCa 4, and perhaps more broadly for TTSs, are dominated by ignoring spots and are not caused by fundamental problems in either photospheric models or in spectral type-effective temperature scales.

@@

@@

One conceivable interpretation of the 67\% to 83\% coverage fraction of cool photosphere is a stellar axis inclined towards our observing direction, with a large circum-polar spot covering much of the observable hemisphere, as described by \citet{donati14}.  The ambient hot photosphere would then be close to the equator, which would be limb-darkened due to projection effects.  Such limb darkening would cause a non-linear transformation between our observed fill factor $f_{\Omega}$, and the surface areal coverage fraction of spots.  So it's conceivable, and probably likely that the reported $f_{\Omega}$ is and overestimate of the surface area actually covered by spots.  In this circum-polar spot scenario, local fingers of cool spots could reach down to near the equator, entering into and out-of the field of view and giving rise to the large amplitude photometric variation seen on LkCa 4.  {\bf Donati-- How does this compare to the reconstructed maps from ZDI in Donati et al. 2014?}.

@@

\section{Conclusions}
{\bf TBD}

@@

%\todo{Verify that the final term possesses a $q$ and not a $q^2$ term.}
